\TheTitle

### 2019-02-20 Paper1/introduction.tex

Numerical methods for approximating the moment $\bm{\hat u_i}$ can be divided into intrusive and non-intrusive methods. The main idea of intrusive methods is to derive a system of equations for the moments and implementing a numerical solver for this system: Testing the initial problem \eqref{eq:hyperbolicProblem} with $\varphi_i$ for $|i|\leq M$ yields
\begin{subequations}\label{eq:nonClosedMomentSystem}
\begin{align}
\partial_t \bm{\hat u}_i(t,\bm{x}) + \nabla&\cdot\langle\bm{F}(\bm{u}(t,\bm{x},\cdot)) \varphi_i\rangle = \bm{0}, \\
\bm{\hat u_i}(t=0,\bm{x}&) = \langle\bm{u}_{\text{IC}}(\bm{x},\cdot)\varphi_i\rangle.
\end{align}
\end{subequations}
To obtain a closed set of equations, one needs to derive a closure $\mathcal{U}$ such that 
\begin{align*}
\bm{u}(t,\bm x,\bm \xi) \approx \mathcal{U}(\bm{\hat u}_0,\cdots,\bm{\hat u}_N;\xi).
\end{align*}
A commonly used closure is the stochastic-Galerkin, which represents the solution by a polynomial:
\begin{align*}
\mathcal{U}_{\text{SG}}(\bm{\hat u}_0,\cdots,\bm{\hat u}_N;\xi):= \sum_{i=0}^N \bm{\hat{u}}_i\varphi_i(\bm{\xi}).
\end{align*}
When using the stochastic-Galerkin method to close \eqref{eq:nonClosedMomentSystem}, the resulting moment system is not necessarily hyperbolic, meaning that it cannot be solved with standard methods. An approach to derive a closure, which ensures hyperbolicity is the Intrusive Polynomial Moment (IPM) closure, which has been presented in \cite{poette2009uncertainty}: The closure is given by a constraint optimization problem. For a given convex entropy $s:\mathbb{R}^p\to\mathbb{R}$ for the original problem \eqref{eq:hyperbolicProblem}, this optimization problem is
\begin{align}\label{eq:primalProblem}
\mathcal{U}_{ME}(\bm{u}) = \argmin_{\bm u} \langle s(\bm u) \rangle \enskip \text{ subject to } \bm{\hat u}_i = \langle \bm u \varphi_i \rangle \text{ for } i = 0,\cdots,N.
\end{align}
This problem can be rewritten as
\begin{align}\label{eq:dualProblem}
 \bm{\hat \lambda}(\bm{\hat u}) := \argmin_{\bm{\lambda} \in \mathbb{R}^{s\cdot(N + 1)}}
  \left\{\langle s_*(\sum_{i=0}^{N}\bm{\lambda}_i \varphi_i)\rangle - \sum_{i=0}^{N}\bm{\lambda}_i^T \bm{\hat u}\right\},
\end{align}
where $s_*:\mathbb{R}\to\mathbb{R}$ is the Legendre transformation of $s$, and $\bm{ \hat\lambda}$ are called the dual variables. The solution to \eqref{eq:primalProblem} is then given by
\begin{align}\label{eq:ansatz}
 \mathcal{U}_{ME}(\bm{u}) = \left( s_{\bm{u}} \right)^{-1}(\bm{\hat{\lambda}}(\bm{u})^T \bm{\varphi}).
\end{align}
The IPM method has several advantages: Choosing the entropy $s(\bm{u}) = \frac{1}{2}\bm{u}^T\bm{u}$ yields the stochastic-Galerkin method, i.e. IPM generalizes different intrusive method. Furthermore, at least for scalar problems, IPM is less oscillatory compared to SG \cite{kusch2017maximum}.

>>>>

Numerical methods for approximating the moment $\bm{\hat u_i}$ can be divided into intrusive and non-intrusive methods. The main idea of intrusive methods is to derive a system of equations for the moments and implementing a numerical solver for this system: Testing the initial problem \eqref{eq:hyperbolicProblem} with $\varphi_i$ for $|i|\leq M$ yields
\begin{subequations}\label{eq:nonClosedMomentSystem}
\begin{align}
\partial_t \bm{\hat u}_i(t,\bm{x}) + \nabla&\cdot\langle\bm{F}(\bm{u}(t,\bm{x},\cdot)) \varphi_i\rangle = \bm{0}, \\
\bm{\hat u_i}(t=0,\bm{x}&) = \langle\bm{u}_{\text{IC}}(\bm{x},\cdot)\varphi_i\rangle.
\end{align}
\end{subequations}
To obtain a closed set of equations, one needs to derive a closure $\mathcal{U}$ such that 
\begin{align*}
\bm{u}(t,\bm x,\bm \xi) \approx \mathcal{U}(\bm{\hat u}_0,\cdots,\bm{\hat u}_N;\xi).
\end{align*}
A commonly used closure is the stochastic-Galerkin, which represents the solution by a polynomial:
\begin{align*}
\mathcal{U}_{\text{SG}}(\bm{\hat u}_0,\cdots,\bm{\hat u}_N;\xi):= \sum_{i=0}^N \bm{\hat{u}}_i\varphi_i(\bm{\xi}).
\end{align*}
When using the stochastic-Galerkin method to close \eqref{eq:nonClosedMomentSystem}, the resulting moment system is not necessarily hyperbolic, meaning that it cannot be solved with standard methods. An approach to derive a closure, which ensures hyperbolicity is the Intrusive Polynomial Moment (IPM) closure, which has been presented in \cite{poette2009uncertainty}: The closure is given by a constraint optimization problem. For a given convex entropy $s:\mathbb{R}^p\to\mathbb{R}$ for the original problem \eqref{eq:hyperbolicProblem}, this optimization problem is
\begin{align}\label{eq:primalProblem}
\mathcal{U}_{ME}(\bm{u}) = \argmin_{\bm u} \langle s(\bm u) \rangle \enskip \text{ subject to } \bm{\hat u}_i = \langle \bm u \varphi_i \rangle \text{ for } i = 0,\cdots,N.
\end{align}
This problem can be rewritten as
\begin{align}\label{eq:dualProblem}
 \bm{\hat \lambda}(\bm{\hat u}) := \argmin_{\bm{\lambda} \in \mathbb{R}^{s\cdot(N + 1)}}
  \left\{\langle s_*(\sum_{i=0}^{N}\bm{\lambda}_i \varphi_i)\rangle - \sum_{i=0}^{N}\bm{\lambda}_i^T \bm{\hat u}_i\right\},
\end{align}
where $s_*:\mathbb{R}^p\to\mathbb{R}$ is the Legendre transformation of $s$, and $\bm{ \hat\lambda}$ are called the dual variables. The solution to \eqref{eq:primalProblem} is then given by
\begin{align}\label{eq:ansatz}
 \mathcal{U}_{ME}(\bm{\hat u}) = \left( s_{\bm{\hat u}} \right)^{-1}(\bm{\hat{\lambda}}(\bm{u})^T \bm{\varphi}).
\end{align}
The IPM method has several advantages: Choosing the entropy $s(\bm{u}) = \frac{1}{2}\bm{u}^T\bm{u}$ yields the stochastic-Galerkin method, i.e. IPM generalizes different intrusive method. Furthermore, at least for scalar problems, IPM is significantly less oscillatory compared to SG \cite{kusch2017maximum}. Also, as discussed in \cite{poette2009uncertainty}, when choosing $s(\bm u)$ to be a physically correct entropy of the deterministic problem, the IPM solution dissipates the entropy
\begin{align*}
S(\bm{\hat u}) := \langle s( \mathcal{U}_{ME}(\bm{\hat u}))\rangle,
\end{align*}
i.e. the IPM method yields a physically correct entropy solution. This again underlines a weakness of stochastic-Galerkin: If $s(\bm{u}) = \frac{1}{2}\bm{u}^T\bm{u}$ is not a correct entropy of the original problem, the SG method can lead to non-physical solution, which can then lead to a failure of the method. The main weakness of the IPM method is its run time, since it requires the repeated evaluation of \eqref{eq:ansatz}, which involves the computation of the optimization problem \eqref{eq:dualProblem}.

### 2019-02-20 Paper1/main.tex

\author[adressJonas]{Jonas Kusch}
\author[adressJannick]{Jannick Wolters}

\address[adressJonas]{Karlsruhe Institute of Technology, Karlsruhe,
    jonas.kusch@kit.edu}
\address[adressJannick]{Karlsruhe Institute of Technology, Karlsruhe, jannick.wolters@kit.edu}

\begin{abstract}
Standard methods for quantifying the effects of uncertainties in hyperbolic problems can yield oscillatory solution approximations. ...
\end{abstract}

@@

\end{frontmatter}
\section{Introduction}

\section{Methods of Uncertainty Quantification}
\subsection{Collocation}
\subsection{Stochastic Galerkin}
\subsection{Intrusive Polynomial Moment Method}
\section{Method discussion}
\begin{itemize}
	\item kleinere Fehler [11]
	\item Programmieraufwand [31]
	\item zweiabhängig
	\item adaptiv (breack up black box approach )
	\item grobe Auflösung für Momente 
\end{itemize}
\section{UQCreator}
\subsection{Adaptivity}
\subsection{restart IPM} %relay(-strategy) IPM?
\subsection{one-shot IPM}
\subsection{adaptive IPM}
\section{Results}
\subsection{2D Euler flow over NACA0012}
\begin{itemize}
	\item AoA study
	\item AoA + Ma number + mehr ?
\end{itemize}
\subsection{2D shallow water dam break}

>>>>

\author[adressJonas]{Jonas Kusch}
\author[adressJannick]{Jannick Wolters}
\author[adressMartin]{Martin Frank}

\address[adressJonas]{Karlsruhe Institute of Technology, Karlsruhe,
    jonas.kusch@kit.edu}
\address[adressJannick]{Karlsruhe Institute of Technology, Karlsruhe, jannick.wolters@kit.edu}
\address[adressMartin]{Karlsruhe Institute of Technology, Karlsruhe, martin.frank@kit.edu}

\begin{abstract}
Methods for quantifying the effects of uncertainties in hyperbolic problems can be divided into intrusive and non-intrusive techniques. Intrusive methods require the implementation of a new code and yield

\end{abstract}

@@

\end{frontmatter}

\input{introduction}
\input{framework}
\input{colloc_ipm}
\input{one_shot_ipm}
\input{results}

%\section{Methods of Uncertainty Quantification}
%\subsection{Collocation}
%\subsection{Stochastic Galerkin}
%\subsection{Intrusive Polynomial Moment Method}
%\section{Method discussion}
%\begin{itemize}
%	\item kleinere Fehler [11]
%	\item Programmieraufwand [31]
%	\item zweiabhängig
%	\item adaptiv (breack up black box approach )
%	\item grobe Auflösung für Momente 
%\end{itemize}
%\section{UQCreator}
%\subsection{Adaptivity}
%\subsection{restart IPM} %relay(-strategy) IPM?
%\subsection{one-shot IPM}
%\subsection{adaptive IPM}
%\section{Results}
%\subsection{2D Euler flow over NACA0012}
%\begin{itemize}
%	\item AoA study
%	\item AoA + Ma number + mehr ?
%\end{itemize}
%\subsection{2D shallow water dam break}
